Fitting the Undetected: Light-curve Feature Estimation in the Presence of Non-detections

### 2013-03-26 tex/ms.tex
new, reduced-scope abstract

\begin{abstract}
In multi-epoch imaging surveys, faint variable sources will not be
detectable at all epochs.  In principle, the best approach is to
perform photometric measurements anyway, so that there is a datum
recorded at every epoch.  In practice, most surveys only provide
detection upper limits at some epochs; in some not even upper
limits but just the information that the star was not observed.  Here
we demonstrate with real data on periodic variables from the All-Sky
Automated Survey (ASAS) that these non-detections are nonetheless useful in
fitting and parameter estimation; their inclusion improves both
precision and accuracy.  One novel aspect of this work is that we do
not require the data source to include accurate---or even
any---information about the uncertainty variances on detections or the
values of the upper limits; we find that we can model these as latent
variables, even when they are different at every epoch.  Using realistic
simulations of Mira variable stars, we obtain amplitude estimates that
are {\bf XXX\%} more accurate by incorporating the non-detections into
the analysis.  For  Mira and RR Lyrae variables observed by ASAS, our
method obtains a tighter period - amplitude relationship than the standard
technique.
\end{abstract}

>>>>

\begin{abstract}
In multi-epoch imaging surveys, faint variable sources will not be
detectable at all epochs.  In principle, the best approach is to
perform photometric measurements (forced photometry) anyway.
%
In practice, most surveys only provide detection upper limits at some
epochs; in some not even upper limits but just the ``null''
information that the star was not detected.
%
Here we demonstrate with real data on periodic variables from the
All-Sky Automated Survey (ASAS) that these null observations can
in fact be very useful in fitting and parameter estimation.
%
We build a novel light-curve model that includes the per-epoch
detection thresholds as a set of latent parameters, one per data
point; the best-fit model is the maximum-likelihood model after
marginalizing out the latent thresholds.
%
The key assumption is that the detection thresholds are drawn from a
stationary distribution (the parameters of which are free to vary).
%
We demonstrate that this censored-data model outperforms traditional
fitting---in which null observations are discarded from the
fits---in estimating both periods and amplitudes for stars subject to
heavy censoring.
%
In particular, our model eradicates a failure mode in which
traditional fitting obtains an integer fraction of the correct period.
\end{abstract}
